Continous time Markov chain models in phylogenetics and phylogeography; model extension, simulation and visualisation

### 2014-06-04 PhDThesis.tex
added examples subsections, they need some work

%TODO: GTR as example? see M USYB 4

@@

% TODO: better illustration, perhaps distances under some model of evolution?
As a simple illustration of the concept of Bayesian analysis let us consider a toy example where two sequences are compared and $x=20$ differences are found out of $N=1000$ sites compared.
Suppose that the parameter $\theta$ we want to infer is the proportion of sites which differ between the two sequences.
Assuming that the sites are independent we can use a binomial model and we put a beta prior with parameters $\alpha=2$, $\beta=100$ on the unknown proportion $\theta$.
Figure \ref{fig:bayes1} shows the resulting posterior distribution along with prior and scaled likelihood.

@@

@@

% metropolisHastings <- function(loglikelihood, prior, proposal, startvalue, data, Nsim) {
% chain = array(dim = c(Nsim, 1))
% chain[1, ] = startvalue
% for (t in 1 : (Nsim - 1)) {
% candidate = proposal(chain[t, ])
% probab = exp( loglikelihood(candidate, data) + prior(candidate, chain[t, ]) - loglikelihood(chain[t, ], data) -
% prior(chain[t, ], candidate) )
% if (runif(1) < probab) {
% chain[t + 1, ] = candidate
% } else {
% chain[t + 1, ] = chain[t, ]
% }#END: accept check
% }#END: iterations loop
% return(chain)
% }

>>>>

%TODO: JK69 as example
% \subsection{Example: JK69 substitution model}

@@

@@

% TODO: better illustration, perhaps distances under some model of evolution?
\subsection{Example}

As a simple illustration of the concept of Bayesian analysis let us consider a toy example where two sequences are compared and $x=20$ differences are found out of $N=1000$ sites compared.
Suppose that the parameter $\theta$ we want to infer is the proportion of sites which differ between the two sequences.
Assuming that the sites are independent we can use a binomial model and we put a beta prior with parameters $\alpha=2$, $\beta=100$ on the unknown proportion $\theta$.

Figure \ref{fig:bayes1} shows the resulting posterior distribution along with prior and scaled likelihood.

@@

% Common choice is to sample from a continuous uniform distribution centered around the current state of the chain: 
% $$\theta^{*}\sim U\left[\theta[t]-\omega,\theta[t]+\omega\right],$$
% where $\omega$ is the arbitrary window size.

\begin{algorithm}
\begin{center}
\begin{algorithmic}[1]
\footnotesize{
%
\State Pick a starting value of the chain $\theta \left[ t \right]$ for $t=0$.
%
\For{int $\left(t=0; \; t<\text{chain length}; \; t++\right)$}
%
\State Generate a candidate state from the proposal distribution (kernel):
$$\theta^{*} \sim q(\theta[t])$$
%
\State Calculate the acceptance ratio:
$$r=\frac{prior\left(\theta^{*}\right)}{prior\left(\theta\left[t\right]\right)}\cdot\frac{Likelihood\left(\mathbf{X}|\theta^{*}\right)}{Likelihood\left(\mathbf{X}|\theta\left[t\right]\right)}.$$
%
\State Sample $u\sim U[0,1]$.
%
\If{$\left( u < r \right)$}
%
\State Set $\theta[t+1]=\theta^*.$
%
\Else 
%
\State Set $\theta[t+1]=\theta[t].$
%
\EndIf
%
\EndFor
}
\end{algorithmic}
\end{center}
\caption{{ \footnotesize {\bf AAA} 
}}
\end{algorithm}

% TODO: better title
\subsection{Example}
